---
title: "p8105_hw3_YZ5290"
output: github_document
date: "2025-10-13"
---

Problem 1
```{r}
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

```{r}
library(p8105.datasets)
data("instacart")
```
```{r}
instacart
```
The dataset contains 138461 observations and 15 variables. Each observation is a single product from an Instacart order. Key variables include order_id: a different ID for each order; product_id: a different ID for each product and product_name which is the name of the product. The example product 	
Bulgarian Yogurt has unique product id 49302 belong to yogurt aisel is in order id 1 and has reordered once. 

```{r}
instacart|>
  count(aisle, sort = TRUE)
```
There are 134 aisle in total and the most items ordered from fresh vergetables. 
```{r}
instacart |>
  count(aisle, sort = TRUE) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = n, y = aisle)) +
  geom_col(fill = "skyblue") +
  labs(
    title = "Number of Items Ordered by Aisle (>10,000 orders)",
    x = "Number of Items Ordered",
    y = "Aisle"
  )
```

```{r}
top3tble=
  instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  count(aisle, product_name, name = "num_orders") |>
  group_by(aisle) |>
  slice_max(order_by = num_orders, n = 3) |>
  ungroup()
top3tble

```








